\textcolor{red}{Simple} connectome inference using partial correlation from imaging calcium signal

### 2014-06-17 article.tex
fix notation
image format pdf

% BG : brain, complex organ
The human brain is a complex biological organ, formed by 100
billions of neurons with 7000 synaptic connections on average.
Cutting edge optical devices delivers in real-time  2-D  images of
the neuronal activity of thousands of neurons through a fluorescent
calcium indicator [todo cite]. Automatic and analytical methods are required to infer
the connectomes from those time series and have to handle several
experimental issues. Some neurons of the network might not be
observed or are confounded with others. The lower device sampling rate with
respect to neuronal activity speed might not allow to distinguish neuron
activation order, especially in highly synchronous activity
[Stetter and in Stetter [20,46,47]] of a large part of the network. Moreover
the fluorescent concentration is highly reactive to a neuron activation or a
spike, but have a slow decay.

% Introduce notation + goal in mathematical score
The connectomes, the neural network, is represented by a (un)directed graph $G = (V, E)$ comprising
a set of nodes $V$ and a set of edges $E \subseteq \left\{(i, j) \in V \times
V\right\}$, (un)ordered pair from $V$.
Each node $i \in V$ represents a neuron, and each edge $(i, j) \in E$
represents a direct link and interactions between neuron $i$ and a neuron $j$.
In the directed graph, each edges also expresses a causal relationship: an
edge $(i, j) \in E$ indicates that the state of neuron $j$ might be caused
by the neuron $i$. The network inference task is set as follows :
\textit{Given the observations $(x^t \in \mathcal{R}^{|V|})_{t=1}^T $
along $T$ periods of time of the $|V|$ neurons, the goal is to infer the
adjacency matrix $A : a_{i,j} = \mathbb{1}((i, j) \in E)$ of the neural network
of graph $G$, where $\mathbb{1}$ is the indicator function.}

@@

Given the image acquisition process, the fluorescent calcium time series
are smoothed using several linear and non linear filters in
order to reduce the imaging process effects. The overall filtering approach
is displayed at Figure \ref{fig:fitlering}.

\begin{figure}[bth]
\includegraphics[width=\linewidth]{images/fig_filtering_recut.png}
\caption{Filtering steps. \textcolor{red}{TODO remake in pdf}}
\label{fig:fitlering}
\end{figure}

The raw signal (original signal of Figure \ref{fig:fitlering}) is very noisy
due to light scattering artifacts that ordinarily affects the quality of the
recording (\cite{lichtman2011big}). A low-pass filter is first applied to
filter high frequency noise with one of the following designs
\begin{align}
% symmetrical median filter
x^t_i &\leftarrow x^{t-1}_i + x^{t}_i + x^{t+1}_i \label{eq:symetric-median}, \\
% asymmetrical weighted median
x^t_i &\leftarrow 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_{t}^i.
\label{eq:weighted-asymetric-median}.
\end{align}

>>>>

% BG : brain, complex organ
The human brain is a complex biological organ, formed by 100
billions of neurons with 7000 synaptic connections on average.
Cutting edge optical devices have been developed to track in real-time
the neuronal activity of thousands of neurons through a fluorescent
calcium indicator  [todo cite]. Automatic and analytical methods are required to infer
the connectomes from those time series.

% Introduce notation + goal in mathematical score
% TODO check convention for y_{ij}
The connectomes, the neural network, is represented by a (un)directed graph $G = (V, E)$ comprising
a set of nodes $V$ and a set of edges $E \subseteq \left\{(i, j) \in V \times
V\right\}$, (un)ordered pair from $V$.
Each node $i \in V$ represents a neuron, and each edge $(i, j) \in E$
represents a direct link and interactions between neuron $i$ and a neuron $j$.
In the directed graph, each edges also expresses a causal relationship: an
edge $(i, j) \in E$ indicates that the state of neuron $j$ might be caused
by the neuron $i$. The network inference task is set as follows :
\textit{Given the observations $(x^t \in \mathcal{R}^{|V|})_{t=1}^T $
along $T$ periods of time of the $|V|$ neurons, the goal is to infer the
adjacency matrix $A : a_{i,j} = \mathbb{1}((i, j) \in E)$ of the neural network
of graph $G$, where $\mathbb{1}$ is the indicator function.}

@@

% \textcolor{red}{This section should be rewritten once we get the figure
                % which displayed the effect of the different filters.}

% Defects of the imaging device
% and will have to cope with
% a slower sampling rate than the neuron firing speed, a superposition
% of the activity of several neurons and the slow decay of fluorescent calcium
% indicator.

The raw signal is the calcium fluorescence imaging representing neuron
activity via a capture imaging process. Observing the fluorescent calcium as
an indicator of neuron activity delivers a 2-D view of neuron network at each
capture which only gives a network screenshot at given time points. It implies
obvious issues. Some neurons may be invisible because they are not observed or
they are confounded with another one because of the loss of the third
dimension. The capture sample rate is also a constraint that could make
difficult to detect in which order the neurons fire.  Moreover the activity
indicator is highly reactive to a burst but takes some time to recover its non
active state.  Finally, the difficulty to distinguish direct relationships when
 a network burst happens, defined as a highly synchronous activity
 [Stetter and in Stetter [20,46,47]], is increased by the slow sampling rate.

Considering the described imaging process, we try in this
section to cope with some of these issues and, in particular, the slower
sampling rate than the neuron firing frequency and the slow decay of
fluorescent calcium indicator after a neuron activity burst.

\begin{figure}%{r}{0.7\linewidth}
\caption{Filtering steps.}\label{wrap-fig:1}
\includegraphics[width=\linewidth]{images/fig_filtering.pdf}
\end{figure}

% Explain why we need to filter

Prior inferring the network, the fluorescent calcium time series
are smoothed using several linear and non linear filters in
order to reduce the imaging process effects. We successively apply different
filters listed below.

As it can be seen on \textit{fig 1}, the raw signal is very noisy due to light
scattering artifacts that ordinarily affects the quality of the recording [in
Stetter, [6]]. Therefore, a low-pass filter is first applied to get a cleaner
version of the signal where only true variations are kept, i.e. without noise
variations. However, the choice of a low-pass filter is not unique. In the
rest of the paper, we will use two particular filters giving similar results,
the symmetrical median filter, i.e., $x^t_i \leftarrow x^{t-1}_i + x^{t}_i +
x^{t+1}_i$, and an asymmetrical weighted median filter taking into account the
history, i.e., $x^t_i \leftarrow 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i
+ x^{t}_i$.
